Model Checking for Open Concurrent Systems

### 2018-09-06 2018-ModelCheckingPNets/checking.tex
More explanation

\newcommand{\A}{\mathcal{A}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\T}{\mathcal{T}}

\tikzset{
	sshadow/.style={opacity=.25, shadow xshift=0.05, shadow yshift=-0.06},
}

%-----#1 size, #2 angle, #3 aspect, #4 label next to the diamond
%-----#5 name of the node, #6 coordinate, #7 label
\def\schemel[#1,#2,#3,#4,#5,#6]#7{ %
	\node[draw, diamond, shape aspect=#3, rotate=#2, minimum size=#1, %
	bottom color=green!55, top color=green!25, color=green!65!black, %
	drop shadow={sshadow,color=green!60!black}, #4] (#5) at #6
	{\textcolor{green!40}{bla}}; %
	\node at #6 {#7};%
}

\def\schemer[#1,#2,#3,#4,#5,#6]#7{ %
	\node[draw, diamond, shape aspect=#3, rotate=#2, minimum size=#1, %
	bottom color=green!65, top color=green!30, color=green!60!black, %
	drop shadow={sshadow,color=green!65!black}, #4] (#5) at #6
	{\textcolor{green!53}{bla}}; %
	\node at #6 {#7}; %
}

%-----TBoxes
%-----#1 height, #2 width, #3 anchor for the label, #4 name of the node, #5
%-----coordinate, #6 label
\def\tboxl[#1,#2,#3,#4,#5]#6{%
	\node[draw, drop shadow={opacity=.35}, minimum height=#1, minimum width=#2, %
	inner color=blue!45, outer color=blue!55, color=blue!40!black] (#4) at #5 {}; %
	\node[anchor=#3,inner sep=2pt] at (#4.#3) {#6};%
}

\def\tboxr[#1,#2,#3,#4,#5]#6{%
	\node[draw, drop shadow={opacity=.35}, minimum height=#1, minimum width=#2, %
	inner color=blue!35, outer color=blue!45, color=blue!50!black] (#4) at #5 {}; %
	\node[anchor=#3,inner sep=2pt] at (#4.#3) {#6}; %
}

%-----#1 name of the node, #2 coordinate, #3 label
\def\entity[#1,#2]#3;{
	\node[draw,drop shadow={opacity=.4,shadow xshift=0.04, shadow
		yshift=-0.04},color=blue!30!black,fill=white,rounded corners=3] (#1) at #2 {#3};
}

%-----#1 from node, #2 to node, #3 specification of a node (label), #4
%-----dashed, or other parameters for draw
\def\isaedge[#1,#2,#3,#4];{ 
	\draw[-triangle 60,color=black!20!black,#4,fill=white] (#1) -- #3
	(#2);  
}

%-----ABoxes
%-----#1 height, #2 width, #3 aspect, #4 name of the node, #5
%-----coordinate, #6 label
\def\aboxl[#1,#2,#3,#4,#5]#6{%
	\node[draw, cylinder, alias=cyl, shape border rotate=90, aspect=#3, %
	minimum height=#1, minimum width=#2, outer sep=-0.5\pgflinewidth, %
	color=orange!40!black, left color=orange!70, right color=orange!80, middle
	color=white] (#4) at #5 {};%
	\node at #5 {#6};%
	\fill [orange!30] let \p1 = ($(cyl.before top)!0.5!(cyl.after top)$), \p2 =
	(cyl.top), \p3 = (cyl.before top), \n1={veclen(\x3-\x1,\y3-\y1)},
	\n2={veclen(\x2-\x1,\y2-\y1)} in (\p1) ellipse (\n1 and \n2); }

\def\aboxr[#1,#2,#3,#4,#5]#6{%
	\node[draw, cylinder, alias=cyl, shape border rotate=90, aspect=#3, %
	minimum height=#1, minimum width=#2, outer sep=-0.5\pgflinewidth, %
	color=orange!50!black, left color=orange!50, right color=orange!60, middle
	color=white] (#4) at #5 {};%
	\node at #5 {#6};%
	\fill [orange!20] let \p1 = ($(cyl.before top)!0.5!(cyl.after top)$), \p2 =
	(cyl.top), \p3 = (cyl.before top), \n1={veclen(\x3-\x1,\y3-\y1)},
	\n2={veclen(\x2-\x1,\y2-\y1)} in (\p1) ellipse (\n1 and \n2); }

%-----#1 height, #2 width, #3 name of the node, #4
%-----coordinate, #5 label
\def\kbbox[#1,#2,#3,#4,#5]#6{
	\draw[dashed] node[draw,color=gray!50,minimum
	height=#1,minimum width=#2] (#4) at #5 {}; 
	\node[anchor=#3,inner sep=2pt] at (#4.#3)  {#6};
}

%-----#1 from node, #2 to node, #3 specification of a node (label), #4
%-----dashed, or other parameters for draw
\def\soledge[#1,#2,#3,#4];{
	\draw[dashed,-latex,#4] (#1) -- #3 (#2);
}

@@

\urldef{\mailstu}\path|{tengfei.li}@stu.ecnu.edu.cn| 
\urldef{\mailecnu}\path|{jliu, hysun}@sei.ecnu.edu.cn|      
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\institute{
%\mailstu\\
%\mailecnu\\
}

@@

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the pM$\mu$} 
	\label{mcl}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula 	&     $\alpha ::= \{c(arg_{1}, ..., arg_{n})\mid arg_{i}=?x\ |\ !e\ |\ \tau \}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

	State formula 	&  $\varphi ::= p \ |\ \neg\varphi \ |\ \varphi_{1}\wedge\varphi_{2}\ |\ \varphi_{1}\vee\varphi_{2} \ |\ [\alpha]\varphi \ |\ \langle\alpha\rangle\varphi$ \\ 
	& $\exists x_{i}\varphi\ |\ \forall x_{i}\varphi\ |\ \mu X(e_{i}/x_{i})\varphi\ |\ \nu X(e_{i}/x_{i})\varphi$	\\ \hline 
	\end{tabular}

\end{table}

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, or a receive variable $?x$. $p$ is an atomic proposition.  
  \TODO{Tengfei: Must complete the explaination}.

We employ the case study from timer component\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by pM$\mu$:

@@

At first, we can do $\langle start?M\rangle$ and then enters a loop.

where, 
\begin{itemize}
\item either it can do a $timeout$ and then we are done and the result is true.
\item or we can do a $tick$ and then we continue the loop, decreasingly, the value t 
\end{itemize}

@@

Modal graph is present by Lin\cite{lin2001modal}. We consider a variant of modal graph, parameterised modal graph (pMG), as the graphical version of pM$\mu$. A pMG is a directed graph $\langle N, E, L_{N}, n_{0}\rangle$, where
\begin{itemize}
	\item $N$ is a set of nodes.
	\item $E \subset N\times N$ is the set of edges.
	\item $L_{N} = (L_{V}, L_{O}, L_{D})$ associates each node $n\in N$ a set of variables $L_{V}(n)$, an operator $L_{O}(n)\in\{\wedge, \vee, \forall x, \exists x,  {\color{red}{\alpha(args)}}, \theta\}\cup  {\color{red}{ModOp}} \cup BExp$, and a depth $L_{D}(n)$ which is a natural number.
	\item $n_{0} \in N$ is the root of pMG.
\end{itemize}

Fig.~\ref{architectrue} presents the architecture of model checking. After we describe the property of a pLTS with pM$\mu$, we transfer the pM$\mu$ formula to pMG. After that, we will {\color{red}{verify the pLTS model with the pMG by the algorithm}}.

@@

The transformation rules can be seen in Lin's paper~\cite{lin2001modal}. The rules take a pM$\mu$ formula as the input and generate a pMG.

\begin{figure}
	\begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=2cm,
	noname/.style={ellipse, minimum width=2.5em, minimum height=1.5em, draw}]
	\node[noname] (1)                                            				 	{$t_{1}$};
	\node[noname] (2) [right=of 1]                               				 	{$t_{2}$};
	\node[noname] (3) [right=of 2]                              				 	{$t_{3}$};
	\node[noname] (4) [node distance=0.5cm and 2cm, above right=of 3]    {$t_{4}$};
	\node[noname] (5) [right=of 4]                                					{$t_{5}$};
	\node[noname] (6) [node distance=0.5cm and 2cm, below right=of 3]     {$t_{6}$};

\section{The algorithm}

@@

 In \cite{lin1996stga}, Lin presents a symbolic transition graph with assignment (STGA) to model concurrent processes. In a STGA, a node is associated with a set of variables. and the edge between nodes is labelled by guard, assignment and action with a tuple $(b, \bar{x}:=\bar{e}, \alpha)$. The first-order $\mu$-calculus is employed as the property language. In order to check the graph, the first-order $\mu$-calculus is transferred to a graphical version through the proposed rules, and then we check {\color{red}{STGA and the modal graph}} by the algorithm.

@@

The difference in parameterised action results in the difference in modality of the algorithm. Based on similarity, we check the pLTS through modifying the algorithm, especially in the modality.

\subsubsection{Alternation-free pM$\mu$}

A formula is said to be alternation-free, when the alternation depth is one, without mutually recursive greatest and least fixed-point operators. A formula expressed with pM$\mu$ is alternation-free. But the formula with first-order $\mu$-calculus have to deal with alternation depth.{\color{red}{[To do Tengfei: definition of alternation depth]}}

@@

{\color{red}{
To to Eric:
In order to construct the abstract data domains, we divide the value of $Max$ to variable and constant. If Max is a variable, Max can express the interval on integer. The change of the value is controlled by a self-decreasing operator. Also, we need another integer 0 as the minimal boundary. If Max is a constant, i.e. Max is the maximal boundary and 0 is the minimal boundary. So we need another variable $pos$ as the interval between Max and 0. Fig.~\ref{side:a} and Fig.~\ref{side:b} are two interpretations on the data domain of the variables. Fig.~\ref{side:a} shows an interpretation that includes two constant (Max, 0), a variable (pos) and an operator (-) on the integer. While Fig.~\ref{side:b} specifies that $Max$ is a variable.
}}

@@

\subsubsection{Processing parameterised action}
We change the algorithm to make it suitable for checking pLTS. From the syntax of pM$\mu$, we define the parameters in the action as $args$, and $args$ can be one of $!e$, $?x$ and $null$ or the composition of them. In the modification version, we change the modality $[c!e], \langle c!e\rangle$, $[c?x], \langle c?x\rangle$ to $[c]$, $\langle c\rangle$, $[c(args)], \langle c(args)\rangle$.
It was shown as follows:

@@

{\color{red}{[The principle of model checking]}}
Given a pLTS model $M$ and a pMG formula $f$, the algorithm checks whether the model satisfies the formula $M\models f$. Each state is made up with a couple $\{s\rho, n\rho'\}$, where $s$ is a state in pLTS, and $n$ a node of pMG. A transition of the model checking graph is from $\{s\rho, n\rho_{0}\}$ to $\{s'\rho', n'\rho_{0}'\}$.

@@

In the initial state, there is no variable for the pLTS. After the action $start(?M)$ triggers the transition from $s_{0}$ to $s_{1}$, the value of the variable $t$ will be changed to $Max$. In the transition, the action expression of the operator in the edge of pMG will trigger the occurrence of transitions in pLTS. The change of value in $s$ depends on the semantics of pLTS. In the state $s_{1}$, the system will check whether the guard $t>0$ satisfies. If yes, the action $tick$ will lead to the occurrence of loop transition. Or, the system will enter another transition back to the initial state $s_{0}$. {\color{red}{[After executing the model checking graph, it will finally return to state $S_{40}$ or state $S_{41}$. Both of them are true.}}

@@

	\bibitem{lin1996stga} Lin, Huimin. "Symbolic transition graph with assignment." International Conference on Concurrency Theory. Springer, Berlin, Heidelberg, 1996.
	\bibitem{cousot1976static}Cousot, Patrick, and Radhia Cousot. "Static determination of dynamic properties of programs." Proceedings of the 2nd International Symposium on Programming, Paris, France. Dunod, 1976.

>>>>

@@

\urldef{\mailtengfei}\path|{tengfeili2018}@gmail.com| 
\urldef{\mailinria}\path|{eric.madelaine}@inria.fr|      
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\institute{
%\mailtengfei\\
%\mailinria\\
}

@@

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the pM$\mu$} 
	\label{mcl}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula 	&     $\alpha ::= \{\{c(arg_{1}, ..., arg_{n}), \tau\}\mid arg_{i}=?x\ |\ !e\  \}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

	State formula 	&  $\varphi ::= p \mid X \mid \neg\varphi \mid \varphi_{1}\wedge\varphi_{2}\ |\ \varphi_{1}\vee\varphi_{2} \ |\ [\alpha]\varphi \ |\ \langle\alpha\rangle\varphi$ \\ 
	& $\exists x_{i}\varphi\ |\ \forall x_{i}\varphi\ |\ \mu X(e_{i}/x_{i})\varphi\ |\ \nu X(e_{i}/x_{i})\varphi$	\\ \hline 
	\end{tabular} 
\end{table}

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, or a receive variable $?x$. An action $\alpha$ is a parameterised action $c(arg_{i})$ or a silent action $\tau$. $p$ is an atomic proposition and $X$ is a set of variables. $\neg$, $\vee$ and $\wedge$ are respectively negation, disjunction and conjunction operator on action formula and state formula. $[\alpha]\varphi$ expresses "whenever an action $\alpha$ happens, we will do $\varphi$", and $\langle\alpha\rangle$ expresses "an action $\alpha$ takes effective, we will do $\varphi$". $x_{i}\in X$ is a free variable bound in state formula $\varphi$. $\mu$ and $\nu$ are respectively the minimum fixed point and the maximum fixed point.\TODO{Tengfei: Must more explaination}

We employ the case study from timer component\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by pM$\mu$ formula:

@@

At first, we can do $\langle start?M\rangle$ and then it enters a loop. After that, either we can do a $timeout$, and then it is done and the result is true. Or we can do a $tick$ and then we continue the loop with decrementing the value $t$ by one.

@@

Modal graph is present by Lin\cite{lin2001modal}. We consider a variant of modal graph, parameterised modal graph (pMG), as the graphical version of pM$\mu$. We define a collection of modal operators as $ModOp = \{[\alpha], \langle\alpha\rangle\}$. A pMG is a directed graph $\langle N, E, L_{N}, n_{0}\rangle$, where
\begin{itemize}
	\item $N$ is a set of nodes.
	\item $E \subseteq N\times N$ is the set of edges.
	\item $L_{N} = (L_{V}, L_{O}, L_{D})$ associates each node $n\in N$ a set of variables $L_{V}(n)$, an operator $L_{O}(n)\in\{\wedge, \vee, \forall x, \exists x, \theta\}\cup ModOp \cup BExp$, and a depth $L_{D}(n)$ which is a natural number.
	\item $n_{0} \in N$ is the root of a pMG.
\end{itemize}

 Given a pM$\mu$ formula $F$ and a pMG, we define a node as a subformula $\phi$ of $F$. The edge between two nodes will be the transition from a subformula to another subformula. The transformation rules can be seen in Lin's paper~\cite{lin2001modal}. The rules take a pM$\mu$ formula as the input and generate a pMG.

\begin{figure}
\centering
	\begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=1.2cm,
	noname/.style={ellipse, minimum width=1.5em, minimum height=1.5em, draw}]
	\node[noname] (1)                                            				 	{$t_{1}$};
	\node[noname] (2) [right=of 1]                               				 	{$t_{2}$};
	\node[noname] (3) [right=of 2]                              				 	{$t_{3}$};
	\node[noname] (4) [node distance=0.5cm and 2cm, above right=of 3]    {$t_{4}$};
	\node[noname] (5) [right=of 4]                                					{$t_{5}$};
	\node[noname] (6) [node distance=0.5cm and 2cm, below right=of 3]     {$t_{6}$};

Fig.~\ref{architectrue} presents the architecture of model checking. The core of the figure is the model checking algorithm, which will be present in Sec.~\ref{algorithm}. The algorithm takes a pLTS model and a pM$\mu$ formula as the input. First, we describe the property of a concurrent system with pM$\mu$ formula and transfer the pM$\mu$ formula to a pMG. And then, the pMG will be verified with the pLTS model by the algorithm. After that, the result will return true or false.

@@

\section{The algorithm}\label{algorithm}

@@

 In \cite{lin1996stga}, Lin presents a symbolic transition graph with assignment (STGA) to model concurrent processes. In a STGA, a node is associated with a set of variables. and the edge between nodes is labelled by guard, assignment and action with a tuple $(b, \bar{x}:=\bar{e}, \alpha)$. The first-order $\mu$-calculus is employed as the property language. In order to check the graph, the first-order $\mu$-calculus is transferred to a graphical version through the proposed rules, and then STGA is checked by the modal graph by the algorithm.

@@

The difference in parameterised action results in the difference in modality of the algorithm. Based on similarity, we have to change the modality and check the pLTS through the modified algorithm.

The notation of alternation depth is given by Emerson and Lei in~\cite{emerson1986alternation} as follows:

\textbf{Definition of Alternation Depth}
Given a pM$\mu$ formula $\Phi$, $\Phi_{i}$ is a subformula of $\Phi$ and $\Gamma_{i}$ is a containing fixed point subformula of $\Phi$. The alternation depth $ad(\Phi)$ of $\Phi$ is defined as follows:

\begin{itemize}
\item[-] $ad(A) = ad(X) = 0$, for any atomic proposition $A$ and variable $X$.
\item[-] $ad(\Phi_{1}\wedge\Phi_{2}) = ad(\Phi_{1}\vee\Phi_{2}) = max\{ ad(\Phi_{1}), ad(\Phi_{2})\}$.
\item[-] $ad([\alpha]\Phi) = ad(\langle\alpha\rangle\Phi) = ad(\Phi)$.
\item[-] Let $\sigma\in\{\mu,\nu\}$, and $\bar{\sigma}$ be the dual of $\sigma$. Then
	\begin{center}
	$ad(\sigma X.\Phi) = max\{1, ad(\Phi), 1+ad(\bar{\sigma} X_{1}.\Phi_{1}), ... , 1+ ad(\bar{\sigma} X_{n}.\Phi_{n})\}$
	\end{center} 
	where $\bar{\sigma} X_{1}.\Phi_{1}, ... , \bar{\sigma} X_{n}.\Phi_{n}$ are subpromula of $\Phi$.
\end{itemize}

A formula is said to be alternation-free, when the alternation depth is one, without mutually recursive greatest and least fixed-point operators. A formula expressed with pM$\mu$ is alternation-free. But the formula with first-order $\mu$-calculus have to deal with alternation depth.\TODO{definition of alternation depth]}
\NOTE{A formula expressed with pM$\mu$ is alternation-free. !!!NO}

@@

\TODO{
In order to construct the abstract data domains, we divide the value of $Max$ to variable and constant. If Max is a variable, Max can express the interval on integer. The change of the value is controlled by a self-decreasing operator. Also, we need another integer 0 as the minimal boundary. If Max is a constant, i.e. Max is the maximal boundary and 0 is the minimal boundary. So we need another variable $pos$ as the interval between Max and 0. Fig.~\ref{side:a} and Fig.~\ref{side:b} are two interpretations on the data domain of the variables. Fig.~\ref{side:a} shows an interpretation that includes two constant (Max, 0), a variable (pos) and an operator (-) on the integer. While Fig.~\ref{side:b} specifies that $Max$ is a variable.
}

@@

\subsubsection{Processing parameterised action}
We change the algorithm to make it suitable for checking pLTS. From the syntax of pM$\mu$, we define the parameters in the action as $args$, and $args$ can be one of $!e$, $?x$ and $null$ or the composition of them. In the modification version, we change the modality from $[c!e], \langle c!e\rangle$, $[c?x], \langle c?x\rangle$ to $[c]$, $\langle c\rangle$, $[c(args)], \langle c(args)\rangle$.
It was shown as follows:

@@

\TODO{The principle of model checking}
Given a pLTS model $M$ and a pM$\mu$ $f$, the algorithm checks whether the model satisfies the formula $M\models f$. First, the pM$\mu$ is transferred to a pMG. As for the compositional model checking graph of pMG and pLTS, each state is made up with a couple $\{s\rho, n\rho'\}$, where $s$ is a state in pLTS, and $n$ a node of pMG. A transition of the model checking graph is from $\{s\rho, n\rho_{0}\}$ to $\{s'\rho', n'\rho_{0}'\}$.  The label of the transition is the same as that of pLTS, and the label of the state is the same as that of the node in pMG.

@@

In the initial state, there is no variable for the pLTS. After the action $start(?M)$ triggers the transition from $s_{0}$ to $s_{1}$, the value of the variable $t$ will be changed to $Max$. In the transition, the action expression of the operator in the edge of pMG will trigger the occurrence of transitions in pLTS. The change of value in $s$ depends on the semantics of pLTS. In the state $s_{1}$, the system will check whether the guard $t>0$ satisfies. If yes, the action $tick$ will lead to the occurrence of loop transition. Or, the system will enter another transition back to the initial state $s_{0}$. \TODO{After executing the model checking graph, it will finally return to state $S_{40}$ or state $S_{41}$. Both of them are true.}

@@

	\bibitem{lin1996stga} Lin, Huimin. "Symbolic transition graph with assignment." International Conference on Concurrency Theory. Springer, Berlin, Heidelberg, 1996.
	\bibitem{emerson1986alternation}Emerson, E. Allen and Chin-Laung Lei. ?Efficient Model Checking in Fragments of the Propositional Mu-Calculus (Extended Abstract).? LICS (1986).
	\bibitem{cousot1976static}Cousot, Patrick, and Radhia Cousot. "Static determination of dynamic properties of programs." Proceedings of the 2nd International Symposium on Programming, Paris, France. Dunod, 1976.
